Array

### 2014-01-04 DOCUMENTATION/AdvancedTypes.tex
Adding support/explanation about the different types of Array %- Declaring the type of array  a : (INT X INT) --> INT %- Declaring a array dynamically {1|->2|->3, 2|->3|->4}

\VGM{To add in text: The nodes \B{Int\_lit} and \B{Bool\_lit} are literals in its 
respective set.}
\section{The LLVM grammar with support to array multidimensional}

@@

Some examples of array are:
\begin{itemize}
	\item $array1 \in (0..10) \to INT $
	\item $array2 \in (0..10) \times (0..10) \to INT $   (a multidimensional example)
	\item $array3 \in BOOL \times (0..10) \to INT$ (a domain with two index types)
\end{itemize}

@@

@@

\begin{ccode} 
  int x=2; int y=3 ; int z=4 ;
  Array3[x][y][z]=7;
\end{ccode}

@@

\paragraph{Restriction on array and structure:} It is not yet supported the 
translation of direct attribution between array expression, for example:

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{pascal}

MACHINE Rec
SETS
 ID={aa,bb}
VARIABLES xx, bank, total
INVARIANT
 xx : struct(name : ID, balance : NAT) & bank : struct(name : ID) &
 total:struct(money:NAT)
INITIALISATION 
 bank := rec(name:bb) ||
 xx:= rec(name:aa, balance:10) ||
 total := rec(money:20)
OPERATIONS

@@

  unsafe_dec = BEGIN
       xx'balance := xx'balance - 1
  END
END

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{pascal}
IMPLEMENTATION
    Rec_i

@@

   bank := rec(name:bb) ;
   xx := rec(name:aa, balance:10) ;
   total :=rec(money:20)
OPERATIONS

  res<--positive = 
  IF xx'balance > 0 THEN
       res:= TRUE 
  ELSE res := FALSE
  END;

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{pascal}

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{pascal}
IMPLEMENTATION 
   Array_i
REFINES
   Array
CONCRETE_VARIABLES
   arr
INVARIANT 
    arr :  ((0..99) --> 0..1000)
INITIALISATION
   arr(1) := 100
OPERATIONS
   set ( ix , tt ) =
   IF
       ix >= 0 & ix <=99 & tt >= 0  & tt <= 1000
   THEN

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{pascal}
MACHINE
   Bubble
VARIABLES
   vec,sort
INVARIANT
   vec : 0..99 --> 0..99 & sort : 0..1 &
   ( sort = 1 =>  !ii.(ii : 0..98 => vec(ii)<=vec(ii+1)))
INITIALISATION
   vec :: 0..99 --> 0..99 || sort := 0
OPERATIONS
   op_sort =
   ANY sorted_vector WHERE
    sorted_vector : 0..99 --> 0..99 &
    !ii.(ii : 0..98 => sorted_vector(ii)<=sorted_vector(ii+1))
   THEN
    vec := sorted_vector|| sort := 1
   END

>>>>

\VGM{To add in text: The nodes \B{Int\_lit} and \B{Bool\_lit} are literals in its 
respective set.}
\VGM{The LLVM grammar with support to array multidimensional.}

@@

Some examples of array are:
\begin{itemize}
	\item $array1 \in (0..10) \to INT $	(a simple array)
	\item $array2 \in (0..10) \times (0..10) \to INT $   (a multidimensional array)
	\item $array3 \in BOOL \times (0..10) \to INT$ (an array with two index types)
	\item $\{1\vdash2\vdash3, 2\vdash3\vdash4\}  \in (INT \times INT) \times INT$ (an explicit array)
\end{itemize}

@@

%TODO: Explain the other rules of Array 
%- Declaring the type of array  a : (INT X INT) --> INT
%- Declaring a array dynamically {1|->2|->3, 2|->3|->4}

@@

\begin{ccode} 
  int x=2; int y=3; int z=4;
  Array3[x][y][z]=7;
\end{ccode}

@@

\paragraph{Restriction on array and structure:} The translation of direct attribution
 between array expression is not yet supported, for example:

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{c}
MACHINE Rec
SETS
 ID={aa,bb}
VARIABLES xx, bank, total
INVARIANT
 xx : struct(name : ID, balance : NAT) & bank : struct(name : ID) &
 total:struct(money:NAT)
INITIALISATION 
 bank := rec(name:bb) ||
 xx:= rec(name:aa, balance:10) ||
 total := rec(money:20)
OPERATIONS

@@

  unsafe_dec = 
  BEGIN
       xx'balance := xx'balance - 1
  END
END
\end{minted}

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{c}
IMPLEMENTATION
    Rec_i

@@

   bank := rec(name:bb) ;
   xx := rec(name:aa, balance:10) ;
   total :=rec(money:20)
OPERATIONS
  res<--positive = 
  IF xx'balance > 0 THEN
       res:= TRUE 
  ELSE res := FALSE
  END;

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{c}

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{c}
IMPLEMENTATION 
   Array_i
REFINES
   Array
CONCRETE_VARIABLES
   arr
INVARIANT 
    arr :  ((0..99) --> 0..1000)
INITIALISATION
   arr(1) := 100
OPERATIONS
   set ( ix , tt ) =
   IF
       ix >= 0 & ix <=99 & tt >= 0  & tt <= 1000
   THEN

@@

\begin{minted}[frame=single,bgcolor=bg,linenos=true,fontsize=\scriptsize]{c}
MACHINE
   Bubble
VARIABLES
   vec,sort
INVARIANT
   vec : 0..99 --> 0..99 & sort : 0..1 &
   ( sort = 1 =>  !ii.(ii : 0..98 => vec(ii)<=vec(ii+1)))
INITIALISATION
   vec :: 0..99 --> 0..99 || sort := 0
OPERATIONS
   op_sort =
   ANY sorted_vector WHERE
    sorted_vector : 0..99 --> 0..99 &
    !ii.(ii : 0..98 => sorted_vector(ii)<=sorted_vector(ii+1))
   THEN
    vec := sorted_vector|| sort := 1
   END
